---
title: "Danish Kings W10 Assignment"
author: "Nikolaj Riemer 202206222"
date: "2025-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)
The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an  `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the separator: Dataene i kings.csv er separeret af kommaer  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=FALSE}
# FILL IN THE CODE BELOW and review the outputs
kings1 <- read.csv("data/kings.csv")

kings2 <- read_csv("data/kings.csv")

kings3 <- read.csv2("data/kings.csv")

kings4 <- read_csv2( "data/kings.csv")
```
Answer: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object
Tidyverse funktionerne er dem med underscore: 'read_csv' og 'read_csv2'. Fordi mine data er kommasepareret læser jeg dem ind som et objekt med funktionen 'read_csv' nedenfor.
2. What is the result of running `class()` on the `kings` object created with a tidyverse function.
Ved at køre funktionen 'class(kings)' får jeg output: [1] "spec_tbl_df" "tbl_df" "tbl" "data.frame" 
Fordi R forstår, at mit data er numerisk, og vi her har indlæst pakken tidyverse, kan den fortælle, at dataene i 'kings' er i tibble-format.
3. How many columns does the object have when created with these different functions?
Funktionen ncol(kings) fortæller mig, at der er 5 kolonner i arket.
4. Show the dataset so that we can see how R interprets each column
R forstår kolonnerne som <dbl>
```{r}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE, then turn the 'eval' flag in this chunk to TRUE.
library(tidyverse)
kings <- read_csv("data/kongerækken.csv")

class(kings)


glimpse(kings)


ncol(kings)


```


## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:


* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`


Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.


```{r}
# YOUR CODE
kingsduration_uden_na <- kings %>% 
  filter(!is.na(sidste_regeringsår))

kingsduration <- kingsduration_uden_na %>%
  mutate(duration= sidste_regeringsår - første_regeringsår)

# Her har jeg fundet varigheden af regeringstiden for danske monarker og samlet dem i en ny kolonne med navnet 'duration'.
```


## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

* You first need to know how to select the relevant `duration` column. What are your options?
* Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
# YOUR CODE
kings_average <- mean(kingsduration$duration)

# Her har jeg udregnet den gennemsnitlige regeringstid for alle de danske monarker, i værdien 'kings_average' = 20
```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in  `kings` dataset. Display the result and also  count the resulting rows with `count()`

```{r}
# YOUR CODE
longer_duration <- kingsduration %>% 
  filter(duration>kings_average)

count(longer_duration)

# 25 danske monarker nød en længere-end-gennemsnitlig regeringstid
```


## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
# YOUR CODE
kings_longest_duration <- kingsduration %>% 
  select(duration, regent) %>% 
  arrange(desc(duration)) %>% 
  slice_max(order_by = duration, n=3)

regeringstid_dage <- kings_longest_duration %>% 
  mutate(regeringstid_dage=duration*365)

regeringstid_dage
  
# Svar: De tre danske monarker med længest regeringstid, regerede i hhv. 21900, 18980 og 15695 dage.  
```


## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
* Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# YOUR CODE
kings_graph <- kingsduration %>%
  mutate(midyear = første_regeringsår + (sidste_regeringsår - første_regeringsår / 2))

kings_graph %>% 
  ggplot(kings_graph, mapping = aes(x = midyear, y = duration))+
  geom_point()+
  geom_smooth()
# Se grafen

```



